---
title: "Genre Trees - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        - text: Corresponding Analysis script
          icon: file-code
          href: "https://github.com/markusradke/GenreDiscourseAnalysis/blob/main/analysis/01_build_genre_trees.R"
        - text: Helper functions - Tree Building
          icon: file-code
          href: "https://github.com/markusradke/GenreDiscourseAnalysis/blob/main/R/build_genre_tree.R"
        - text: Helper functions - Initial Genre Mapping
          icon: file-code
          href: "https://github.com/markusradke/GenreDiscourseAnalysis/blob/main/R/map_genres.R"
        - text: Helper functions - Tree Plotting
          icon: file-code
          href: "https://github.com/markusradke/GenreDiscourseAnalysis/blob/main/R/make_network_graph.R"
        css: 
          styles.css
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()
```	

This report presents interactive genre networks for the open music database MusicBrainz and the digital music streaming platform Spotify. All visualizations are based on the POPTRAG dataset, which contains a large collection of tracks that were listened to in Germany within the last 50 years and their corresponding genre tag assignements on both plattforms.

# The intital genre network
In the initial genre network, **each node represents a genre tag** on the corresponding platform. The **edges of this initial network are based on the co-occurrence** of genre tags across all tracks in the POPTRAG dataset (i.e., the probability that a certain genre tag appears, given that another genre tag was assigned to a track; see example with "pop" and "hip hop" below). The more frequently two genres co-occur, the stronger the connection (edge) between them. Musicbrainz provides additional community votes for genre assignments (for Spotify we used multiple artists per track to generate votes), which are used to **weight the edges genre network**. Here, we multiplied the link strengh by the mean relative vote for a genre across all tracks where both genres appear. The more users voted for a genre, i.e. the more unambigously a track can be assigned to a genre, the stronger the connection becomes (see continued example).



```{mermaid}
flowchart LR
rock -- P(metal|rock) * meanRelFreqVotes(metal|rock) --> metal
metal -- P(rock|metal) * meanRelFreqVotes(rock|metal) --> rock
```

# Network of subgenres
Next, we determine which genre is a subgerne of the other. Genres with subgenres should be distributed more evenly across multiple genres. Thus, in our initial genre network, we eliminate the weaker connection between the two genres. To account for the uncertainty in this decision, we subtract the weaker conditional probability from the conditional probability of the stronger connection (see the continued example). The resulting directed network represents a **genre genealogy**, where edges point from subgenres to their parent genres.

```{mermaid}
flowchart LR
rock -. P(metal|rock) .-x metal
metal -- P(rock|metal) - P(rock|metal) --> rock
```


# 'Radically democratic' genre trees
Now, we transform the directed network of descendant genres into a hierarchical tree. We assume that the strongest link to a potential supergenre determines the most suitable parent genre. Thus, we eliminate all other outgoing links to potential supergenres (see the continued example).

```{mermaid}
flowchart LR
metal --> rock
metal -.-x pop
```

# Prune tree to improve structure
Finally, we remove edges with very low weight (= weak parenthood). Corresponding genres are reattached to the root node of the tree. This step is tuned to  minimize the weighted Gini accross hiearrchy levels, ensuring a more balanced distribution of genres across the tree structure.

# Genres Trees 
## Genre Tree for MusicBrainz

### Tree Building Statistics
```{r musicbrainz-tree-stats, echo=FALSE}
get_degree_stats <- function(graph){
  in_degree <- igraph::degree(graph, mode = "in")
  out_degree <- igraph::degree(graph, mode = "out")
  list(
    in_degree_median = median(in_degree),
    in_degree_IQR = IQR(in_degree),
    in_degree_min = min(in_degree),
    in_degree_max = max(in_degree),
    node_in_degree_mins = names(which(in_degree == min(in_degree))),
    node_in_degree_maxs = names(which(in_degree == max(in_degree))),
    out_degree_median = median(out_degree),
    out_degree_IQR = IQR(out_degree), 
    out_degree_min = min(out_degree),
    out_degree_max = max(out_degree), 
    node_out_degree_mins = names(which(out_degree == min(out_degree))),
    node_out_degree_maxs = names(which(out_degree == max(out_degree)))
    )
}

initial_network <- readRDS("../models/trees/MusicBrainz_initial_network.rds")
bidirectional_removed <- readRDS("../models/trees/MusicBrainz_bidirectional_removed.rds")
tree_graph <- readRDS("../models/trees/MusicBrainz_graph.rds")

no_components_initial <- igraph::components(initial_network)$no
density_initial <- igraph::edge_density(initial_network)
no_nodes_initial <- igraph::vcount(initial_network)
degree_stats_initial <- get_degree_stats(initial_network)
degree_stats_bidirectional_removed <- get_degree_stats(bidirectional_removed)
degree_stats_tree <- get_degree_stats(tree_graph)
```	

The tree comprises `r no_nodes_initial` genres (nodes). The initial genre network had `r no_components_initial` disconnected components and a density of `r round(density_initial, 4)*100`% (observed connections / potential connections in the network). 

```{r musicbrainz-degree-table, echo=FALSE}
render_degree_stages_table <- function(
  degree_stats_initial,
  degree_stats_bidirectional_removed,
  degree_stats_tree,
  platform_name) { 
  degree_table <- data.frame(
    Metric = c("In-Degree Median", "In-Degree IQR", "In-Degree Min", "In-Degree Max",
              "Out-Degree Median", "Out-Degree IQR",  "Out-Degree Min", "Out-Degree Max"),
    Initial_Network = c(degree_stats_initial$in_degree_median,
                        degree_stats_initial$in_degree_IQR,
                        degree_stats_initial$in_degree_min,
                        degree_stats_initial$in_degree_max,
                        degree_stats_initial$out_degree_median,
                        degree_stats_initial$out_degree_IQR,
                        degree_stats_initial$out_degree_min,
                        degree_stats_initial$out_degree_max),
    Bidirectional_Removed = c(degree_stats_bidirectional_removed$in_degree_median,
                              degree_stats_bidirectional_removed$in_degree_IQR,
                              degree_stats_bidirectional_removed$in_degree_min,
                              degree_stats_bidirectional_removed$in_degree_max,
                              degree_stats_bidirectional_removed$out_degree_median,
                              degree_stats_bidirectional_removed$out_degree_IQR,
                              degree_stats_bidirectional_removed$out_degree_min,
                              degree_stats_bidirectional_removed$out_degree_max),
    Final_Tree = c(degree_stats_tree$in_degree_median,
                  degree_stats_tree$in_degree_IQR,
                  degree_stats_tree$in_degree_min,
                  degree_stats_tree$in_degree_max,
                  degree_stats_tree$out_degree_median,
                  degree_stats_tree$out_degree_IQR,
                  degree_stats_tree$out_degree_min,
                  degree_stats_tree$out_degree_max)
  )
  knitr::kable(degree_table, caption = sprintf("Degree Statistics Across Tree Building Stages for %s Genre Tree", platform_name))
}
render_degree_stages_table(
  degree_stats_initial,
  degree_stats_bidirectional_removed,
  degree_stats_tree,
  platform_name = "MusicBrainz"
)
```
**Initial Network**<br>
Nodes with minimal number of connections:<br>*`r degree_stats_initial$node_in_degree_mins`*<br>
Nodes with maximum number of connections:<br>*`r degree_stats_initial$node_out_degree_maxs`*

**Biderectional Removed Network**<br>
Nodes with minimum in-degree:<br>*`r degree_stats_bidirectional_removed$node_in_degree_mins`* <br>
Nodes with maximum in-degree:<br>*`r degree_stats_bidirectional_removed$node_in_degree_maxs`* <br>
Nodes with minimum out-degree:<br>*`r degree_stats_bidirectional_removed$node_out_degree_mins`* <br>
Nodes with maximum out-degree:<br>*`r degree_stats_bidirectional_removed$node_out_degree_maxs`*

**Final Tree**<br>
Nodes with maximum in-degree:<br>*`r degree_stats_tree$node_in_degree_maxs`*

### Tree balancing
```{r musicbrainz-tree-pruning, echo=FALSE, fig.width=8, fig.height=6}	
#| fig-cap: "Gini coefficient across iterations of cutting weak genre connections and connecting resulting trees to root"
mb_gini <- readRDS("../models/trees/MusicBrainz_gini_evaluation.rds")
plot <- plot_gini_trajectory(mb_gini)
ggplot2::ggsave(
  "../reports/paper_v1_poetics/figures/balancing_gini_trajectory.png", 
  plot, 
  width = 180, 
  height = 115,
  units = "mm",
  dpi = 300,#300, 
  device = ragg::agg_png)
plot
```	


### Final Tree
:::{.column-screen-inset-right}
:::{.plot-container}
**Navigation:** Click and drag to pan • Mouse wheel to zoom in/out • Use search bar to find and highlight genres (green) • Click blue nodes to expand/collapse

**Controls:** "Center Root" - return to root | "Expand All" - show all genres | "Fold All" - collapse tree
```{r musicbrainz-interactive-tree, echo=FALSE}
mb <- read_feather_with_lists("../data/filtered_mb_long.feather")
plot_interactive_tree_graph(
  readRDS("../models/trees/MusicBrainz_graph.rds"),
  get_sizes_lookup(mb),
  get_fills_lookup(mb), 
  height = 1000
)
```
:::
:::



## Genre Tree for Spotify


### Tree Building Statistics
```{r spotify-tree-stats, echo=FALSE}
initial_network <- readRDS("../models/trees/Spotify_initial_network.rds")
bidirectional_removed <- readRDS("../models/trees/Spotify_bidirectional_removed.rds")
tree_graph <- readRDS("../models/trees/Spotify_graph.rds")

no_components_initial <- igraph::components(initial_network)$no
density_initial <- igraph::edge_density(initial_network)
no_nodes_initial <- igraph::vcount(initial_network)
degree_stats_initial <- get_degree_stats(initial_network)
degree_stats_bidirectional_removed <- get_degree_stats(bidirectional_removed)
degree_stats_tree <- get_degree_stats(tree_graph)
```	

The tree comprises `r no_nodes_initial` genres (nodes). The initial genre network had `r no_components_initial` disconnected components and a density of `r round(density_initial, 4)*100`% (observed connections / potential connections in the network). 

```{r spotify_degree-table, echo=FALSE}
render_degree_stages_table(
  degree_stats_initial,
  degree_stats_bidirectional_removed,
  degree_stats_tree,
  platform_name = "Spotify"
)
```

**Initial Network**<br>
Nodes with minimal number of connections:<br>*`r degree_stats_initial$node_in_degree_mins`*<br>
Nodes with maximum number of connections:<br>*`r degree_stats_initial$node_out_degree_maxs`*

**Biderectional Removed Network**<br>
Nodes with minimum in-degree:<br>*`r degree_stats_bidirectional_removed$node_in_degree_mins`* <br>
Nodes with maximum in-degree:<br>*`r degree_stats_bidirectional_removed$node_in_degree_maxs`* <br>
Nodes with minimum out-degree:<br>*`r degree_stats_bidirectional_removed$node_out_degree_mins`* <br>
Nodes with maximum out-degree:<br>*`r degree_stats_bidirectional_removed$node_out_degree_maxs`*

**Final Tree**<br>
Nodes with maximum in-degree:<br>*`r degree_stats_tree$node_in_degree_maxs`*

### Tree balancing
```{r spotify-tree-balancing, echo=FALSE, fig.width=8, fig.height=6}	
#| fig-cap: "Gini Coefficient Across Tree Pruning Steps"
s_gini <- readRDS("../models/trees/Spotify_gini_evaluation.rds")
plot_gini_trajectory(s_gini)
```	


### Final Tree
Please click on the nodes to explore sub-genres. The font size is fixed to ensure readability. Node size represents the number of songs in each genre.

:::{.column-screen-inset-right}
:::{.plot-container}
**Navigation:** Click and drag to pan • Mouse wheel to zoom in/out • Use search bar to find and highlight genres (green) • Click blue nodes to expand/collapse

**Controls:** "Center Root" - return to root | "Expand All" - show all genres | "Fold All" - collapse tree
```{r spotify-interactive-tree, echo=FALSE}	
s <- read_feather_with_lists("../data/filtered_s_long.feather")
plot_interactive_tree_graph(
  readRDS("../models/trees/Spotify_graph.rds"),
  get_sizes_lookup(s),
  get_fills_lookup(s), 
  height = 1000
)
```
:::
:::


# Towards a Genre Genealogy 
Radically democratic genre trees provide a clear and simple hierarchical representation of genre relationships. However, they also discard potentially valuable information about alternative parent-child relationships between genres. To retain this information, we additionally construct a more comprehensive **genre genealogy network** for qualitative exploration.

For this, we transform the directed network of subgenres into an **Acyclic Directed Graph (ADG)**: We iteratively remove the weakest edges in any remaining cycles. Critically, this cycle-removal process **never removes edges that are part of the genre tree**, ensuring consistency between the genealogy network and the tree representation. The tree is therefore guaranteed to be a subgraph of the genealogy ADG, as it only contains each genre's single strongest parent relationship.


## Genre Genealogy for MusicBrainz
:::{.column-screen-inset-right}
:::{.plot-container}
**Navigation:** Click and drag to pan • Mouse wheel to zoom in/out • Click nodes to change focus • Use search bar to find genres • Adjust edge threshold slider to filter weak connections 

**Green nodes** are part of the genre tree and will always be displayed (no matter which treshold).

**Controls:** "Focus Root" - return to root | "Reset Position" - realign view | "Back" - previous node
```{r musicbrainz-interactive-genealogy, echo=FALSE}
mb <- read_feather_with_lists("../data/filtered_mb_long.feather")
plot_interactive_genealogy_graph(
  readRDS("../models/trees/MusicBrainz_genealogy.rds"),
  readRDS("../models/trees/MusicBrainz_graph.rds"),
  get_sizes_lookup(mb),
  get_fills_lookup(mb),
  height = 1000
)
```
:::
:::

## Genre Genealogy for Spotify
:::{.column-screen-inset-right}
:::{.plot-container}
**Navigation:** Click and drag to pan • Mouse wheel to zoom in/out • Click nodes to change focus • Use search bar to find genres • Adjust edge threshold slider to filter weak connections

**Green nodes** are part of the genre tree and will always be displayed (no matter which treshold).

**Controls:** "Focus Root" - return to root | "Reset Position" - realign view | "Back" - previous node
```{r spotify-interactive-genealogy, echo=FALSE}	
s <- read_feather_with_lists("../data/filtered_s_long.feather")
plot_interactive_genealogy_graph(
  readRDS("../models/trees/Spotify_genealogy.rds"),
  readRDS("../models/trees/Spotify_graph.rds"),
  get_sizes_lookup(s),
  get_fills_lookup(s),
  height = 1000
)
```
:::
:::