---
title: "Metagenre Mapping - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        css: 
          styles.css
---

# Metagenre assignment 
```{r load poptrag, message=FALSE, warning=FALSE}
devtools::load_all()
poptrag <- readRDS("../data/poptrag_selected.rds")

mb_low <- read_feather_with_lists("../models/metagenres/mb_metagenres_low.feather")
mb_medium <- read_feather_with_lists("../models/metagenres/mb_metagenres_medium.feather")
mb_high <- read_feather_with_lists("../models/metagenres/mb_metagenres_high.feather")
mb_very_high <- read_feather_with_lists("../models/metagenres/mb_metagenres_very_high.feather")

s_low <- read_feather_with_lists("../models/metagenres/s_metagenres_low.feather")
s_medium <- read_feather_with_lists("../models/metagenres/s_metagenres_medium.feather")
s_high <- read_feather_with_lists("../models/metagenres/s_metagenres_high.feather")
s_very_high <- read_feather_with_lists("../models/metagenres/s_metagenres_very_high.feather")
```

Once an optimal metagenre configuration has been selected based on the tuning results, we can assign metagenres to our songs. For the assignment, all tags that were assigned to a track by platform users were mapped to their corresponding metagenres using the selected solution's mapping table. If a track had multiple tags that mapped to different metagenres, we employed a majority voting approach to determine the final metagenre assignment. As a tie breaker, we selected the metagenre with fewest overall votes assigned to it in the dataset, promoting a more balanced distribution across metagenres.
Please find below the distribution of metagenres in the assigned datasets as well as the median relative probabilities for the assigned metagenres based on the majority votes.

```{r metagenenre descriptives preparement}	
plot_metagenre_prevs <- function(mapping){
  meta_prev <- mapping |> dplyr::count(metagenre, sort = TRUE) |> 
    dplyr::mutate(relfreq = n / sum(n) * 100,
                  metagenre = forcats::fct_inorder(metagenre) |> forcats::fct_rev()) 
    
  ggplot2::ggplot(meta_prev, 
                  ggplot2::aes(x = relfreq, y = metagenre)) +
    ggplot2::geom_bar(stat = "identity", fill = "grey45") +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%s", format(n, big.mark = ","))), 
                     hjust = -0.1, size = 4) +
    ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.2))) +
    ggplot2::labs(y = "",
                  x = "Relative Frequency (%)") +
    ggplot2::theme_minimal() + 
    ggplot2::theme(axis.text.y = ggplot2::element_text(size = 12, face = "bold"), 
                   axis.text.x = ggplot2::element_text(size = 12), 
                   axis.title.x = ggplot2::element_text(size = 12, color = "grey45"), 
                   panel.grid.major.y = ggplot2::element_blank(), 
                   panel.grid.minor.y = ggplot2::element_blank())
}

plot_metagenre_assignement_certainty <- function(mapping){
  prob_cols <- mapping |>
    dplyr::select(dplyr::starts_with("p_"), "metagenre")
  
  max_p <- do.call(pmax, as.list(
    prob_cols |> dplyr::select(-"metagenre")
  ))
  
  meta_probs <- prob_cols |>
    dplyr::mutate(max_p = max_p) |>
    dplyr::group_by(metagenre) |>
    dplyr::summarise(median_probability = median(max_p, na.rm = TRUE), 
                     iqr_probability = IQR(max_p, na.rm = TRUE)) |>
    dplyr::arrange(dplyr::desc(median_probability)) |> 
    dplyr::mutate(metagenre = forcats::fct_inorder(metagenre) |> 
                    forcats::fct_rev())
  
  ggplot2::ggplot(meta_probs, 
                  ggplot2::aes(x = median_probability, y = metagenre)) +
    ggplot2::geom_bar(stat = "identity", fill = "grey65") +
    ggplot2::geom_errorbar(ggplot2::aes(
      xmin = median_probability - iqr_probability / 2, 
      xmax = median_probability + iqr_probability / 2), 
      width = 0.2, color = "black") +
    ggplot2::scale_x_continuous(
      expand = ggplot2::expansion(mult = c(0, 0.05)), 
      breaks = c(0.25, 0.5, 0.75, 1)) +
    ggplot2::labs(y = "",
                  x = "Median Relative Freq of Votes (%)") +
    ggplot2::theme_minimal() + 
    ggplot2::theme(
      axis.text.y = ggplot2::element_text(size = 12, face = "bold"), 
      axis.text.x = ggplot2::element_text(size = 12), 
      axis.title.x = ggplot2::element_text(size = 12, color = "grey45"), 
      panel.grid.major.y = ggplot2::element_blank(), 
      panel.grid.minor.y = ggplot2::element_blank())
}
```

## MusicBrainz Metagenre Assignment
### MusicBrainz `r mb_low$metagenre |> unique() |> length()` Metagenres
```{r mb low prev}
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(mb_low)
```	
```{r mb low certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(mb_low)
```
:::{.column-page-inset-right}
```{r mb low contingency, fig.height=6, fig.width=10}
# | fig-cap: "Contingency table of genre category assignments between GOMGET-G and predicted Rosamerica genres (Essentia) for the low-granularity solution for external validation"
plot <- plot_external_contingency(mb_low, poptrag, "track.ab.genrerosamerica")
ggplot2::ggsave(
  "../reports/paper_v1/figures/rosa_contingency.png", 
  plot, 
  width = 180, 
  height = 160,
  units = "mm"
)
plot
```
:::
### MusicBrainz `r mb_medium$metagenre |> unique() |> length()` Metagenres
```{r mb medium prev}	
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(mb_medium)
```

```{r mb medium certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(mb_medium)
```
:::{.column-page-inset-right}
```{r mb medium contingency,  fig.width=10, fig.height=8.6}
# | fig-cap: "Contingency table of genre category assignments between COMGET-G and Discogs (first mentioned genre) for medium granularity solution for external validation"
plot <- plot_external_contingency(mb_medium, poptrag, "album.dc.firstgenre")
ggplot2::ggsave(
  "../reports/paper_v1/figures/dc_contingency.png", 
  plot, 
  width = 180, 
  height = 160,
  units = "mm"
)
plot
```
:::

### MusicBrainz `r mb_high$metagenre |> unique() |> length()` Metagenres
```{r mb high prev}	
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(mb_high)
```
:::{.column-page-inset-right}
```{r mb high contingency, fig.height=8, fig.width=11}
# | fig-cap: "Contingency table of genre category assignments between COMGET-G and Deezer (first mentioned genre) for the high granularity solution for external validation"
plot <- plot_external_contingency(mb_high, poptrag, "track.dz.album.firstgenre.name")
ggplot2::ggsave(
  "../reports/paper_v1/figures/dz_contingency.png", 
  plot, 
  width = 320, 
  height = 220,
  units = "mm"
)
plot
```
:::
```{r mb high certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(mb_high)
```

### MusicBrainz `r mb_very_high$metagenre |> unique() |> length()` Metagenres
```{r mb very high prev}	
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(mb_very_high)
```

```{r mb very high certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(mb_very_high)
```



## Spotify Metagenre Assignment
### Spotify `r s_low$metagenre |> unique() |> length()` Metagenres
```{r s low prev}
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(s_low)
```
```{r s low certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(s_low)
```
:::{.column-page-inset-right}
```{r s low contingency, fig.height=6, fig.width=10}
# | fig-cap: "Contingency table of genre category assignments between Spotify and predicted Rosamerica genres (Essentia) for the low-granularity solution for external validation"
plot_external_contingency(s_low, poptrag, "track.ab.genrerosamerica", "Spotify")
```
:::
### Spotify `r s_medium$metagenre |> unique() |> length()` Metagenres
```{r s medium prev}
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(s_medium)
```
```{r s medium certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(s_medium)
```
:::{.column-page-inset-right}
```{r s medium contingency, fig.width=10, fig.height=8.6}
# | fig-cap: "Contingency table of genre category assignments between Spotify and Discogs (first mentioned genre) for medium granularity solution for external validation"
plot_external_contingency(s_medium, poptrag, "album.dc.firstgenre", "Spotify")
```
:::

### Spotify `r s_high$metagenre |> unique() |> length()` Metagenres
```{r s high prev}
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(s_high)
```
```{r s high certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(s_high)
```

:::{.column-page-inset-right}
```{r s high contingency, fig.width=10, fig.height=8.6}
# | fig-cap: "Contingency table of genre category assignments between Spotify and Deezer (first mentioned genre) for the high granularity solution for external validation"
plot_external_contingency(s_high, poptrag, "track.dz.album.firstgenre.name", "Spotify")
```
:::


### Spotify `r s_very_high$metagenre |> unique() |> length()` Metagenres
```{r s very high prev}
#| fig-cap: "Assignement prevalences and corresponding number of tracks in the dataset per metagenre"
plot_metagenre_prevs(s_very_high)
```
```{r s very high certainty}
# | fig-cap: "Median relative frequency of votes for assigned metagenres with interquartile range error bars"
plot_metagenre_assignement_certainty(s_very_high)
```

# Metagenre Examples

```{r examples_functions}
top_n <- 5

build_examples <- function(pop, met, top_n) {
  joined_raw <- pop |>
    dplyr::inner_join(
      met |> dplyr::select(track.s.id, p_max, n_max, metagenre),
      by = "track.s.id"
    ) |>
    dplyr::select(
      track.s.title,
      artist.s.name,
      album.s.title,
      album.s.coverurl,
      album.s.releaseyear,
      track.s.previewurl,
      metagenre,
      p_max,
      n_max
    ) |>
    dplyr::filter(!is.na(track.s.previewurl),
                  !is.na(album.s.coverurl))

  joined <- joined_raw |>
    dplyr::mutate(
      track.s.title = sanitize_triple_colon(track.s.title),
      artist.s.name = sanitize_triple_colon(artist.s.name),
      album.s.title = sanitize_triple_colon(album.s.title)
    )

  top <- joined |>
    dplyr::group_by(metagenre) |>
    dplyr::arrange(dplyr::desc(p_max), dplyr::desc(n_max),
                   .by_group = TRUE) |>
    dplyr::distinct(artist.s.name, .keep_all = TRUE) |>
    dplyr::slice_head(n = top_n) |>
    dplyr::ungroup() |>
    dplyr::mutate(selection = "Highest Certainty")

  mid <- joined |>
    dplyr::group_by(metagenre) |>
    dplyr::arrange(dplyr::desc(p_max), dplyr::desc(n_max),
                   .by_group = TRUE) |>
    dplyr::distinct(artist.s.name, .keep_all = TRUE) |>
    dplyr::slice((dplyr::n() %/% 2 - (top_n %/% 2)):
                   (dplyr::n() %/% 2 + (top_n %/% 2))) |>
    dplyr::ungroup() |>
    dplyr::mutate(selection = "Median Certainty")

  low <- joined |>
    dplyr::group_by(metagenre) |>
    dplyr::arrange(p_max, n_max, .by_group = TRUE) |>
    dplyr::distinct(artist.s.name, .keep_all = TRUE) |>
    dplyr::slice_head(n = top_n) |>
    dplyr::arrange(dplyr::desc(p_max), dplyr::desc(n_max),
                   .by_group = TRUE) |>
    dplyr::ungroup() |>
    dplyr::mutate(selection = "Lowest Certainty")

  dplyr::bind_rows(top, mid, low)
}

make_cover_html <- function(url) {
  paste0('<img src="', url, '" style="width:100px;">')
}

make_preview_html <- function(url) {
  paste0('<audio controls style="width:200px;">',
         '<source src="', url, '" type="audio/mpeg"></audio>')
}

sanitize_triple_colon <- function(x) {
  if (is.null(x)) return(x)
  gsub(":::", "&#58;&#58;&#58;", x, fixed = TRUE)
}

make_music_table_df <- function(examples, genre, top_n) {
  examples |>
    dplyr::filter(metagenre == genre) |>
    dplyr::mutate(p_max = round(p_max * 100)) |>
    dplyr::rename(
      Titel = track.s.title,
      Artist = artist.s.name,
      Album = album.s.title,
      Cover = album.s.coverurl,
      Year = album.s.releaseyear,
      Preview = track.s.previewurl,
      "Probability [%]" = p_max,
      Votes = n_max,
      Selection = selection
    ) |>
    dplyr::mutate(
      Titel = gsub("[<>$:\\\"]", "", Titel),
      Artist = gsub("[<>$:\\\"]", "", Artist),
      Album = gsub("[<>$:\\\"]", "", Album),
      Titel = gsub("(?<=\\d)\\.(?=\\s)", "", Titel, perl = TRUE),
      Album = gsub("(?<=\\d)\\.(?=\\s)", "", Album, perl = TRUE),
      Cover = make_cover_html(Cover),
      Preview = make_preview_html(Preview)
    ) |>
    dplyr::select(
      Titel, Artist, Album, Cover, Year,
      Preview, "Probability [%]", Votes, Selection
    )
}

kable_html_for_table <- function(music_table, top_n) {
  k <- knitr::kable(music_table, format = "html", escape = FALSE)
  k <- kableExtra::kable_styling(k, bootstrap_options = "striped",
                                full_width = TRUE, position = "center")
  k <- k |>
    kableExtra::column_spec(column = 1, width = "15%") |>
    kableExtra::column_spec(column = 2, width = "12%") |>
    kableExtra::column_spec(column = 3, width = "12%") |>
    kableExtra::column_spec(column = 4, width = "8%")  |>
    kableExtra::column_spec(column = 5, width = "5%")  |>
    kableExtra::column_spec(column = 6, width = "15%") |>
    kableExtra::column_spec(column = 7, width = "8%")  |>
    kableExtra::column_spec(column = 8, width = "5%")  |>
    kableExtra::column_spec(column = 9, width = "12%") |>
    kableExtra::collapse_rows(columns = 9, valign = "middle") |>
    kableExtra::row_spec(top_n,
                         extra_css = "border-bottom: 3px solid #000 !important;") |>
    kableExtra::row_spec(top_n * 2,
                         extra_css = "border-bottom: 3px solid #000 !important;")
  k
}

render_all_tables <- function(examples, top_n) {
  genres <- unique(examples$metagenre)
  for (g in seq_along(genres)) {
    gen <- genres[g]
    gen_title <- sanitize_triple_colon(gen)
    cat(paste0('### ', gen_title, '\n\n'))
    cat('::: {.column-screen-inset-right}\n\n')
    tbl <- make_music_table_df(examples, gen, top_n)
    k <- kable_html_for_table(tbl, top_n)
    cat(k)
    cat('\n\n:::\n\n')
  }
}
```

## MusicBrainz Low-Detailed Metagenres Examples
```{r mb low examples, results='asis'}	
examples <- build_examples(poptrag, mb_low, top_n)
render_all_tables(examples, top_n)
```

## MusicBrainz Medium-Detailed Metagenres Examples
```{r mb medium examples, results='asis'}
examples <- build_examples(poptrag, mb_medium, top_n)
render_all_tables(examples, top_n)
```

## MusicBrainz High-Detailed Metagenres Examples
```{r mb high examples, results='asis'}
examples <- build_examples(poptrag, mb_high, top_n)
render_all_tables(examples, top_n)
```

## MusicBrainz Very High-Detailed Metagenres Examples
```{r mb very high examples, results='asis'}
examples <- build_examples(poptrag, mb_very_high, top_n)
render_all_tables(examples, top_n)
```

## Spotify Low-Detailed Metagenres Examples
```{r s low examples, results='asis'}
examples <- build_examples(poptrag, s_low, top_n)
render_all_tables(examples, top_n)
```

## Spotify Medium-Detailed Metagenres Examples
```{r s medium examples, results='asis'}
examples <- build_examples(poptrag, s_medium, top_n)
render_all_tables(examples, top_n)
```

## Spotify High-Detailed Metagenres Examples
```{r s high examples, results='asis'}
examples <- build_examples(poptrag, s_high, top_n)
render_all_tables(examples, top_n)
```

## Spotify Very High-Detailed Metagenres Examples
```{r s very high examples, results='asis'}
examples <- build_examples(poptrag, s_very_high, top_n)
render_all_tables(examples, top_n)
```
