---
title: "Random Forest Classification - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        mermaid: 
          theme: neutral
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        css: 
          styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()
data_settings <- readRDS("../models/classifier/data_settings.rds")

train_low <- readRDS("../models/classifier/train_low.rds")
test_low <- readRDS("../models/classifier/test_low.rds")
train_high <- readRDS("../models/classifier/train_high.rds")
test_high <- readRDS("../models/classifier/test_high.rds")

```

This report presents the evaluation results of different classifier models trained to predict the MusicBrainz folksonomy meta-genres. Classifiers were trained for both low detailed (10-15 genres) and high detailed (25-30) meta-genre taxonomies.

The metagenre POPULAR MUSIC was discarded for analysis as the metagenre was only present for too few tracks.


# Complete Workflow
:::{.column-screen-inset-right}
```{mermaid}
%%{init: {
  'flowchart': {
    'useMaxWidth': true,
    'htmlLabels': false
  }
}}%%

flowchart TB
 subgraph Data["Data prep"]
    direction LR
        B["select potential features"]
        A["import POPTRAG"]
        C["transform chars to factors </br> apply lyrics heuristics"]
        D["drop observations with more than 40% missing"]
        E["join metagenre </br>target labels"]
  end
 subgraph Split["Spliting Strategy"]
    direction LR
        F["split train/test <br> (0.8/0.2, artist-level)"]
        H["create 5-fold CV splits </br>(artist-level)"]
  end
 subgraph Pre["Preprocessing"]
    direction LR
        I["impute median/mode"]
        J["downsample majority classes </br>(factor tuned)"]
        K["upsample minority </br>(SMOTE-NC, factor tuned)"]
        N["remove zero-variance predictors"]
        O["normalize numeric </br>(z-score)"]
  end
 subgraph Model["Modeling and eval"]
    direction LR
        P["tune hyperparams on F1 macro </br>(Bayesian optimization, </br>20-set latin hypercube inital grid,</br> max 50 iters, 10-iters stop)"]
        Q["select best params </br>(based on F1 macro)"]
        R["fit final model </br>(on full training set)"]
        S["evaluate on test set"]
  end
    A --> B
    B --> C
    C --> D
    D --> E
    F --> H
    I -. only during training .-> J
    J -.-> K
    K -.-> N
    N --> O
    I --> N
    P --> Q 
    P <--> Pre
    Q --> R
    R --> S 
    R <--> Pre
    Data --> Split
    Split --> Model
    S <--> Pre

    style Pre fill:#f8f0e8,stroke:#cc6600,stroke-width:3px,color:#000
    style Data fill:#e8f4f8,stroke:#0066cc,stroke-width:3px,color:#000
    style Split fill:#f0e8f8,stroke:#6600cc,stroke-width:3px,color:#000
    style Model fill:#e8f8e8,stroke:#00aa00,stroke-width:3px,color:#000
  
```
:::

# Data Summary

```{r dataset_summary}	
summarize_ds <- function(resolution, split, df) {
  data.frame(
    Metagenre_Resolution = resolution,
    Split = split,
    N_Tracks = format(nrow(df), big.mark = ","),
    N_Artists = format(nrow(dplyr::distinct(df, .data$artist.s.id)), big.mark = ",")
  )
}

datasets <- dplyr::bind_rows(
  summarize_ds("low (10-15)",  "Train", train_low),
  summarize_ds("low (10-15)",  "Test",  test_low),
  summarize_ds("high (25-30)", "Train", train_high),
  summarize_ds("high (25-30)", "Test",  test_high),
)

knitr::kable(
  datasets,
  col.names = c("Metagenre Resolution", "Split", "N~Tracks~", "N~Artists~"),
  caption = "Dataset splits used in the analyses"
)
```	

# Split outcome distributions

```{r outcome_distributions}	
# dist_low <- compare_metagenre_distributions(rf_data, detail_level = "low")
# dist_low$plot
# dist_high <- compare_metagenre_distributions(rf_data, detail_level = "high")
# dist_high$plot
```

