---
title: "Metagenre Classification - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        mermaid: 
          theme: neutral
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        css: 
          styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()

data_settings <- readRDS("../models/classifier/data_settings.rds")

train_low <- readRDS("../models/classifier/train_low.rds")
test_low <- readRDS("../models/classifier/test_low.rds")
cv_splits_low <- readRDS("../models/classifier/cv_splits_low.rds")
train_high <- readRDS("../models/classifier/train_high.rds")
test_high <- readRDS("../models/classifier/test_high.rds")
cv_splits_high <- readRDS("../models/classifier/cv_splits_high.rds")

top_n <- 40 # top n models to show in report
n_glmnet_varimp <- 5
n_rf_varimp <- 20

lowres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_10_15.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max)) 

highres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_25_30.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max))

# glmnet_params <- c("penalty", "under_ratio", "over_ratio")
glmnet_params <- c("penalty")
# rf_params <- c("max.depth", "min_n", "mtry", "under_ratio", "over_ratio")
rf_params <- c("max.depth", "min_n", "mtry")

glmnet_col_names <- c(
  "Rank", "Iteration", "Penalty", #"Oversampling Ratio", "Undersampling Ratio",
  "F1 Macro (mean ± se)", "Kappa (mean ± se)", "Accuracy (mean ± se)", ""
)
rf_col_names <- c(
  "Rank", "Iteration", "Max Depth", "Min Node Size", "MTRY",
  # "Oversampling Ratio", "Undersampling Ratio",
  "F1 Macro (mean ± se)", "Kappa (mean ± se)", "Accuracy (mean ± se)", ""
)
```

This report presents the evaluation results of different classifier models trained to predict the MusicBrainz folksonomy meta-genres. Classifiers were trained for both low detailed (10-15 genres) and high detailed (25-30) meta-genre taxonomies.

The metagenre POPULAR MUSIC was discarded for analysis as the metagenre was only present for too few tracks.


# Complete Workflow
:::{.column-screen-inset-right}
```{mermaid}
%%{init: {
  'flowchart': {
    'useMaxWidth': true,
    'htmlLabels': false
  }
}}%%

flowchart TB
 subgraph Data["Data prep"]
    direction LR
        import["import POPTRAG"]
        select["select potential features"]
        transform["transform chars to factors </br> apply lyrics heuristics"]
        casewise["drop observations with more than 40% missing"]
        target["join metagenre </br>target labels"]
        minn["drop levels with less than 100 observations in one of the splits / folds"]
        dummy["dummy code categorical predictors</br>reference level: most common"]
  end
 subgraph Split["Spliting Strategy"]
    direction LR
        testtrain["split train/test <br> (0.8/0.2, artist-level)"]
        cvfold["create 10-fold CV splits </br>(artist-level)"]
  end
 subgraph Pre["Preprocessing"]
    direction LR
        impute["impute median"]
        down["nearmiss downsample majority classes </br>(factor tuned)"]
        up["upsample minority </br>(broderline SMOTE, factor tuned)"]
        rmzv["remove zero-variance predictors"]
        normalize["normalize numeric </br>(z-score)"]
  end
 subgraph Model["Modeling and eval"]
    direction LR
        tune["tune hyperparams on F1 macro </br>(Bayesian optimization, </br>20-set latin hypercube inital grid,</br> max 50 iters, 10-iters stop)"]
        best["select best params </br>(based on F1 macro, one standard error rule)"]
        finalfit["fit final model </br>(on full training set)"]
        evaluate["evaluate on test set"]
  end
    import --> select
    select --> transform
    transform --> casewise
    casewise --> target
    target --> minn
    minn --> dummy
    minn <-.-> Split

    testtrain --> cvfold

    impute -. only during training .-> down
    impute --> rmzv
    down -.-> up
    up -.-> rmzv
    rmzv --> normalize
    tune --> best
    tune <--> Pre
    best --> finalfit
    finalfit --> evaluate
    evaluate <--> Pre
    dummy --> Split
    Split --> Model

    style Pre fill:#f8f0e8,stroke:#cc6600,stroke-width:3px,color:#000
    style Data fill:#e8f4f8,stroke:#0066cc,stroke-width:3px,color:#000
    style Split fill:#f0e8f8,stroke:#6600cc,stroke-width:3px,color:#000
    style Model fill:#e8f8e8,stroke:#00aa00,stroke-width:3px,color:#000
  
```
:::

:::{.callout-caution}
Currently, the models are tranined with a fixed undersampling rate with factor 10 (low-detail) and 20 (high-detail) due to computational constraints. Also, the oversampling rate is fixed to 0.25 for both low-detail and high-detailed metagenre labels.
:::

:::{.callout-note}
The mixture of lasso and ridge regularization for the GLMNET models was fixed to 0.5 (elastic net).

Random forests were tuned with 750 trees per ensemble. The final model on the whole tranining set was trained with 1000 trees.
:::

# Data Summary

```{r dataset_summary}
summarize_ds <- function(resolution, split, df) {
  data.frame(
    Metagenre_Resolution = resolution,
    Split = split,
    N_Tracks = format(nrow(df), big.mark = ","),
    N_Artists = format(nrow(dplyr::distinct(df, .data$artist.s.id)), big.mark = ",")
  )
}

datasets <- dplyr::bind_rows(
  summarize_ds("low (10-15)",  "Train", train_low),
  summarize_ds("low (10-15)",  "Test",  test_low),
  summarize_ds("high (25-30)", "Train", train_high),
  summarize_ds("high (25-30)", "Test",  test_high)
)

knitr::kable(
  datasets,
  col.names = c("Metagenre Resolution", "Split", "N~Tracks~", "N~Artists~"),
  caption = "Dataset splits used in the analyses"
)
```

# Split outcome distributions

```{r outcome_distributions_low}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **low-detailed** meta-genre taxonomy."
distribution_low <- compare_metagenre_distributions_splits(train_low, test_low, cv_splits_low)
distribution_low$plot
```

```{r outcome_distributions_high}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **high-detailed** detailed meta-genre taxonomy."
distribution_high <- compare_metagenre_distributions_splits(train_high, test_high, cv_splits_high)
distribution_high$plot
```

# Random Baseline

This random baseline model predicts the metagenre labels by random guessing according to the distribution of the metagenres in the training set.

:::{.column-screen-inset-right}

```{r random_baseline}
random_low <- load_classification_model("../models/classifier/baseline", "random_baseline_lowres")
random_high <- load_classification_model("../models/classifier/baseline", "random_baseline_highres")
```
Low resolution metrics:
**`r get_metrics_description(random_low$evaluation$metrics_test)`**

High resolution metrics:
**`r get_metrics_description(random_high$evaluation$metrics_test)`**
:::

# Model Features

:::{.column-screen-inset-right}
```{r model_features}
model_features <- colnames(train_low)[
  !colnames(train_low) %in% c("track.s.id", "artist.s.id", "metagenre", "case_wts")
] |> sort()
```


A total of `r length(model_features)` features was used for training the classifiers.

```{r model_features_table}
model_features_padded <- c(
  model_features,
  rep("", times = 5 - (length(model_features) %% 5))
)

knitr::kable(
  matrix(model_features_padded, ncol = 5, byrow = TRUE),
  col.names = c("Model features", rep("", 4))
) |>
  kableExtra::kable_styling(full_width = FALSE, position = "left") |>
  kableExtra::column_spec(1:5, border_right = FALSE, border_left = FALSE) |>
  kableExtra::row_spec(2:ceiling(length(model_features_padded) / 5), hline_after = FALSE)
```
:::

:::{.callout-note}
Variables starting with "dtb." refer to artist genres from Spotify. We assume that artist genres are assigned with a strategic aim for distribution on the platform (Valverde, 2025). In that regard, they are different from the musical genres assigned by the listeners. The variables are flags indicating whether an artist genre is assigned to a track's artist or not. To reduce the dimensionality, we used the same hiearchical-tree-based approach as for the MusicBrainz meta-genres.
:::

# GLMNET Classifier

## Low-detailed metagenre labels

### Tuning

#### Model Selection

```{r glmnet_low_load}
glmnet_low <- load_classification_model("../models/classifier/glmnet", "glmnet_lowres")

glmnet_low_chosen <- c(
  penalty = glmnet_low$settings$penalty#,
  # under_ratio = glmnet_low$settings$under_ratio,
  # over_ratio = glmnet_low$settings$over_ratio
)

glmnet_low_topmodels <- prepare_tuning_topmodels(
  glmnet_low$tuning_history, glmnet_low_chosen, top_n
)
```

```{r glmnet_low_tuning_graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **GLMNET classifier for low-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."
plot_tuning_model_selection(glmnet_low_topmodels, top_n)
```

```{r glmnet_low_tuning_table}
glmnet_low_table <- create_glmnet_top_table(glmnet_low_topmodels)
render_tuning_top_table(glmnet_low_table, glmnet_col_names)
```

#### Parameter Exploration

:::{.column-screen-inset-right}
```{r glmnet_low_tuning_params}
#| layout-ncol: 2
for (i in seq_along(glmnet_params)) {
  print(plot_parameter_tuning(glmnet_low$tuning_history, glmnet_params[i], i == 1))
}
```
:::

### Evaluation on test set

**`r get_metrics_description(glmnet_low$evaluation$metrics_test)`**

**Correlation between metagenre uncertainty and per-class F1 scores on test set: `r round(cor(lowres_certainty$certainty, calculate_class_f1_scores_from_confusion_df(glmnet_low$evaluation$confusion_test)),3)`**

```{r glmnet_low_cm, fig.width=9, fig.height=6}
#| column: page-inset-right
plot_cm(glmnet_low$evaluation, lowres_certainty, "test")
```
### Most important variables per metagenre

```{r glmnet_low_varimp}
render_glmnet_varimp_table(
  glmnet_low$model, n_glmnet_varimp,
  "Top 5 most important features per metagenre according to the GLMNET classifier for low-detailed metagenre labels."
)
```

## High-detailed metagenre labels

### Tuning

#### Model Selection

```{r glmnet_high_load}
glmnet_high <- load_classification_model("../models/classifier/glmnet", "glmnet_highres")

glmnet_high_chosen <- c(
  penalty = glmnet_high$settings$penalty#,
  # under_ratio = glmnet_high$settings$under_ratio,
  # over_ratio = glmnet_high$settings$over_ratio
)

glmnet_high_topmodels <- prepare_tuning_topmodels(
  glmnet_high$tuning_history, glmnet_high_chosen, top_n
)
```

```{r glmnet_high_tuning_graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **GLMNET classifier for high-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."
plot_tuning_model_selection(glmnet_high_topmodels, top_n)
```

```{r glmnet_high_tuning_table}
glmnet_high_table <- create_glmnet_top_table(glmnet_high_topmodels)
render_tuning_top_table(glmnet_high_table, glmnet_col_names)
```

#### Parameter Exploration

:::{.column-screen-inset-right}
```{r glmnet_high_tuning_params}
#| layout-ncol: 2
for (i in seq_along(glmnet_params)) {
  print(plot_parameter_tuning(glmnet_high$tuning_history, glmnet_params[i], i == 1))
}
```
:::

### Evaluation on test set

**`r get_metrics_description(glmnet_high$evaluation$metrics_test)`**

**Correlation between metagenre uncertainty and per-class F1 scores on test set: `r round(cor(highres_certainty$certainty, calculate_class_f1_scores_from_confusion_df(glmnet_high$evaluation$confusion_test)), 3)`**

```{r glmnet_high_cm, fig.width=11.5, fig.height=8.3}
#| column: screen-inset-right
plot_cm(glmnet_high$evaluation, highres_certainty, "test")
```

### Most important variables per metagenre

```{r glmnet_high_varimp}
render_glmnet_varimp_table(
  glmnet_high$model, n_glmnet_varimp,
  "Most important features per metagenre according to the GLMNET classifier for high-detailed metagenre labels."
)
```

# Random Forest Classifier

## Low-detailed metagenre labels

### Tuning

#### Model Selection

```{r rf_low_load}
rf_low <- load_classification_model("../models/classifier/rf", "rf_lowres")

rf_low_chosen <- c(
  max.depth = rf_low$settings$max.depth,
  min_n = rf_low$settings$min_n,
  mtry = rf_low$settings$mtry#,
  # under_ratio = rf_low$settings$under_ratio,
  # over_ratio = rf_low$settings$over_ratio
)

rf_low_topmodels <- prepare_tuning_topmodels(
  rf_low$tuning_history, rf_low_chosen, top_n
)
```

```{r rf_low_tuning_graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **Random Forest classifier for low-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."
plot_tuning_model_selection(rf_low_topmodels, top_n)
```

```{r rf_low_tuning_table}
rf_low_table <- create_rf_top_table(rf_low_topmodels)
render_tuning_top_table(rf_low_table, rf_col_names)
```

#### Parameter Exploration

:::{.column-screen-inset-right}
```{r rf_low_tuning_params}
#| layout-ncol: 2
for (i in seq_along(rf_params)) {
  print(plot_parameter_tuning(rf_low$tuning_history, rf_params[i], i == 1))
}
```
:::

### Evaluation on test set

**`r get_metrics_description(rf_low$evaluation$metrics_test)`**

**Correlation between metagenre uncertainty and per-class F1 scores on test set: `r round(cor(lowres_certainty$certainty, calculate_class_f1_scores_from_confusion_df(rf_low$evaluation$confusion_test)),3)`**
```{r rf_low_cm, fig.width=9, fig.height=6}
#| column: page-inset-right
plot_cm(rf_low$evaluation, lowres_certainty, "test")
```

### Most important variables per metagenre

```{r rf_low_varimp}
plot_varimp_bar(rf_low$evaluation$varimp, n_rf_varimp)
```

## High-detailed metagenre labels

### Tuning

#### Model Selection

```{r rf_high_load}
rf_high <- load_classification_model("../models/classifier/rf", "rf_highres")

rf_high_chosen <- c(
  max.depth = rf_high$settings$max.depth,
  min_n = rf_high$settings$min_n,
  mtry = rf_high$settings$mtry#,
  # under_ratio = rf_high$settings$under_ratio,
  # over_ratio = rf_high$settings$over_ratio
)

rf_high_topmodels <- prepare_tuning_topmodels(
  rf_high$tuning_history, rf_high_chosen, top_n
)
```

```{r rf_high_tuning_graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **Random Forest classifier for high-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."
plot_tuning_model_selection(rf_high_topmodels, top_n)
```

```{r rf_high_tuning_table}
rf_high_table <- create_rf_top_table(rf_high_topmodels)
render_tuning_top_table(rf_high_table, rf_col_names)
```

#### Parameter Exploration

:::{.column-screen-inset-right}
```{r rf_high_tuning_params}
#| layout-ncol: 2
for (i in seq_along(rf_params)) {
  print(plot_parameter_tuning(rf_high$tuning_history, rf_params[i], i == 1))
}
```
:::

### Evaluation on test set

**`r get_metrics_description(rf_high$evaluation$metrics_test)`**

**Correlation between metagenre uncertainty and per-class F1 scores on test set: `r round(cor(highres_certainty$certainty, calculate_class_f1_scores_from_confusion_df(rf_high$evaluation$confusion_test)),3)`**


```{r rf_high_cm, fig.width=11, fig.height=8.3}
#| column: screen-inset-right
plot_cm(rf_high$evaluation, highres_certainty, "test")
```


### Most important variables per metagenre

```{r rf_high_varimp}
plot_varimp_bar(rf_high$evaluation$varimp, n_rf_varimp)
```