---
title: "Metagenre Classification - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        mermaid: 
          theme: neutral
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        css: 
          styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()
data_settings <- readRDS("../models/classifier/data_settings.rds")

train_low <- readRDS("../models/classifier/train_low.rds")
test_low <- readRDS("../models/classifier/test_low.rds")
cv_splits_low <- readRDS("../models/classifier/cv_splits_low.rds")
train_high <- readRDS("../models/classifier/train_high.rds")
test_high <- readRDS("../models/classifier/test_high.rds")
cv_splits_high <- readRDS("../models/classifier/cv_splits_high.rds")

top_n <- 20 # show for tuning
n_glmnet_varimp <- 5
n_rf_varimp <- 20
```

This report presents the evaluation results of different classifier models trained to predict the MusicBrainz folksonomy meta-genres. Classifiers were trained for both low detailed (10-15 genres) and high detailed (25-30) meta-genre taxonomies.

The metagenre POPULAR MUSIC was discarded for analysis as the metagenre was only present for too few tracks.


# Complete Workflow
:::{.column-screen-inset-right}
```{mermaid}
%%{init: {
  'flowchart': {
    'useMaxWidth': true,
    'htmlLabels': false
  }
}}%%

flowchart TB
 subgraph Data["Data prep"]
    direction LR
        import["import POPTRAG"]
        select["select potential features"]
        transform["transform chars to factors </br> apply lyrics heuristics"]
        casewise["drop observations with more than 40% missing"]
        target["join metagenre </br>target labels"]
        minn["drop levels with less than 100 observations in one of the splits / folds"]
        dummy["dummy code categorical predictors</br>reference level: most common"]
  end
 subgraph Split["Spliting Strategy"]
    direction LR
        testtrain["split train/test <br> (0.8/0.2, artist-level)"]
        cvfold["create 5-fold CV splits </br>(artist-level)"]
  end
 subgraph Pre["Preprocessing"]
    direction LR
        impute["impute median"]
        down["nearmiss downsample majority classes </br>(factor tuned)"]
        up["upsample minority </br>(broderline SMOTE, factor tuned)"]
        rmzv["remove zero-variance predictors"]
        normalize["normalize numeric </br>(z-score)"]
  end
 subgraph Model["Modeling and eval"]
    direction LR
        tune["tune hyperparams on F1 macro </br>(Bayesian optimization, </br>20-set latin hypercube inital grid,</br> max 50 iters, 10-iters stop)"]
        best["select best params </br>(based on F1 macro, one standard error rule)"]
        finalfit["fit final model </br>(on full training set)"]
        evaluate["evaluate on test set"]
  end
    import --> select
    select --> transform
    transform --> casewise
    casewise --> target
    target --> minn
    minn --> dummy
    minn <-.-> Split

    testtrain --> cvfold

    impute -. only during training .-> down
    impute --> rmzv
    down -.-> up
    up -.-> rmzv
    rmzv --> normalize
    tune --> best
    tune <--> Pre
    best --> finalfit
    finalfit --> evaluate
    evaluate <--> Pre
    dummy --> Split
    Split --> Model

    style Pre fill:#f8f0e8,stroke:#cc6600,stroke-width:3px,color:#000
    style Data fill:#e8f4f8,stroke:#0066cc,stroke-width:3px,color:#000
    style Split fill:#f0e8f8,stroke:#6600cc,stroke-width:3px,color:#000
    style Model fill:#e8f8e8,stroke:#00aa00,stroke-width:3px,color:#000
  
```
:::

:::{.callout-caution}
Currently, the models are tranined with a fixed undersampling rate with factor 10 (low-detail) and 20 (high-detail) due to computational constraints. Also, the oversampling rate is fixed to 0.25.
:::

:::{.callout-note}
The mixture of Lasso and Ridge regularization for the GLMNET models was fixed to 0.5 (elastic net).

Random forest were tuned with 750 trees per ensemble. The final model on the whole tranining set was trained with 1000 trees.
:::

# Data Summary

```{r dataset_summary}	
summarize_ds <- function(resolution, split, df) {
  data.frame(
    Metagenre_Resolution = resolution,
    Split = split,
    N_Tracks = format(nrow(df), big.mark = ","),
    N_Artists = format(nrow(dplyr::distinct(df, .data$artist.s.id)), big.mark = ",")
  )
}

datasets <- dplyr::bind_rows(
  summarize_ds("low (10-15)",  "Train", train_low),
  summarize_ds("low (10-15)",  "Test",  test_low),
  summarize_ds("high (25-30)", "Train", train_high),
  summarize_ds("high (25-30)", "Test",  test_high),
)

knitr::kable(
  datasets,
  col.names = c("Metagenre Resolution", "Split", "N~Tracks~", "N~Artists~"),
  caption = "Dataset splits used in the analyses"
)
```	

# Split outcome distributions

```{r outcome_distributions low}	
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **low-detailed** meta-genre taxonomy."
distribution_low <- compare_metagenre_distributions_splits(train_low, test_low, cv_splits_low)
distribution_low$plot
```

```{r outcome_distributions high}	
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **high-detailed** detailed meta-genre taxonomy."
distribution_high <- compare_metagenre_distributions_splits(train_high, test_high, cv_splits_high)
distribution_high$plot
```

# Random Baseline
This random baseline model predicts the metagenre labels by random guessing according to the distribution of the metagenres in the training set.

:::{.column-screen-inset-right}

```{r random_baseline low}
#| layout-ncol: 2

random_low <- load_classification_model(
  "../models/classifier/baseline",
  "random_baseline_lowres"
)
plot_cm(random_low$evaluation, "train")
plot_cm(random_low$evaluation, "test")
```	


```{r random_baseline high, fig.width=10, fig.height=8.3}
#| layout-ncol: 2

random_high <- load_classification_model(
  "../models/classifier/baseline",
  "random_baseline_highres"
)
plot_cm(random_high$evaluation, "train")
plot_cm(random_high$evaluation, "test")

```




# Model Features
```{r model_features}	
model_features <- colnames(train_low)[
  !colnames(train_low) %in%
    c(
      "track.s.id",
      "artist.s.id",
      "metagenre",
      "case_wts", 
      "p_max"
    )
] |> sort()
```
A total of `r length(model_features)` features as used for training the classifiers.

```{r model_features_table}

model_features_padded <- c(
  model_features,
  rep("", times = 5 - (length(model_features) %% 5))
)

knitr::kable(
  matrix(
    model_features_padded,
    ncol = 5,
    byrow = TRUE
  ),
  col.names = c("Model features", rep("", 4)),
) |> 
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:5,
    border_right = FALSE,
    border_left = FALSE, 
  ) |> 
  kableExtra::row_spec(
    2:ceiling(length(model_features_padded) / 5),
    hline_after = FALSE
  )

``` 
:::

# GLMENT Classifier
## Low-detailed metagenre labels
### Tuning
#### Model Selection
```{r glmnet low tuning top graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **GLMNET classifier for low-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."

glmnet_low <- load_classification_model(
  "../models/classifier/glmnet",
  "glmnet_lowres"
)


chosen_parameters <- c(
  penalty = glmnet_low$settings$penalty,
  under_ratio = glmnet_low$settings$under_ratio, 
  over_ratio = glmnet_low$settings$over_ratio
)
# find matching model in tuning history
tuning_topmodels <- glmnet_low$tuning_history |> 
  dplyr::mutate(
    is_chosen = dplyr::if_else(
      penalty == chosen_parameters["penalty"] &
        under_ratio == chosen_parameters["under_ratio"] &
        over_ratio == chosen_parameters["over_ratio"],
      TRUE,
      FALSE
    )
  ) |> 
  dplyr::arrange(dplyr::desc(macro_f1_with_zeros_mean)) |> 
  head(top_n) |> 
  dplyr::mutate(rank = dplyr::row_number())
best_f1 <- tuning_topmodels$macro_f1_with_zeros_mean[1]
best_f1_se <- tuning_topmodels$macro_f1_with_zeros_std_err[1]
tuning_topmodels <-tuning_topmodels |> 
  dplyr::mutate(
    within_se_rule = macro_f1_with_zeros_mean >= (best_f1 - best_f1_se), 
    modelgroup = dplyr::case_when(
      is_chosen ~ 1,
      within_se_rule ~ 2,
      TRUE ~ 3
    ) |> as.factor()
  )


ggplot2::ggplot(tuning_topmodels, ggplot2::aes(x = rank, y = macro_f1_with_zeros_mean, color = modelgroup)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = macro_f1_with_zeros_mean - macro_f1_with_zeros_std_err,
      ymax = macro_f1_with_zeros_mean + macro_f1_with_zeros_std_err
    ),
    width = 0.2
  ) +
  ggplot2::labs(
    x = "Rank",
    y = "F1 Macro (mean ± se)"
  ) +
  ggplot2::scale_color_manual(
    values = c("#23cf00ff", "grey25", "grey75"),
  ) +
  ggplot2::scale_x_discrete(position = "top", limits = factor(1:top_n)) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.position = "none", 
    axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold"),
    panel.grid.major.x = ggplot2::element_blank()
  )

```	
```{r glmnet low tuning top table}

top_table <- tuning_topmodels |> 
  dplyr::mutate(
    `F1 Macro (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      macro_f1_with_zeros_mean,
      macro_f1_with_zeros_std_err
    ), 
    `Kappa (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      kap_mean,
      kap_std_err
    ), 
    `Accuracy (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      accuracy_mean,
      accuracy_std_err
    ), 
    `Undersampling Ratio` = round(under_ratio, 2),
    `Oversampling Ratio` = round(over_ratio, 2),
    Penalty = format(penalty, scientific = TRUE, digits = 3)
  ) |> 
    dplyr::select(
      Rank = rank,
      Iteration = .iter,
      Penalty,
      `Oversampling Ratio`,
      `Undersampling Ratio`,
      `F1 Macro (mean ± se)`,
      `Kappa (mean ± se)`,
      `Accuracy (mean ± se)`, 
      modelgroup
    )

knitr::kable(
  top_table,
  col.names = c(
    "Rank",
    "Iteration",
    "Penalty",
    "Oversampling Ratio",
    "Undersampling Ratio",
    "F1 Macro (mean ± se)",
    "Kappa (mean ± se)",
    "Accuracy (mean ± se)",
    ""
  )
) |> 
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:8,
    border_right = FALSE,
    border_left = FALSE, 
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 1),
    bold = TRUE,
    background = "#23cf00ff"
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 2),
    background = "grey90"
  ) |> 
  kableExtra::row_spec(
   which(top_table$modelgroup == 3),
    color = "grey35"
  ) |>
  kableExtra::remove_column(9)
 
```	
#### Parameter Exploration
:::{.column-screen-inset-right}
```{r glmnet low tuning param graphs}
#| layout-ncol: 2
make_axis_title <- function(x) {
  with_space <- gsub("[_\\.]", " ", x)
  split <- strsplit(with_space, " ")[[1]]
  capitalized <- paste(toupper(substring(split, 1, 1)), substring(split, 2),
        sep = "", collapse = " ")
  capitalized
}

plot_parameter_tuning <- function(parameter, plot_legend) {
  ggplot2::ggplot(glmnet_low$tuning_history, ggplot2::aes(x = .data[[parameter]], y = macro_f1_with_zeros_mean, color = .iter)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::scale_x_continuous(position = "top") +
    ggplot2::labs(
      x = make_axis_title(parameter),
      y = "F1 Macro (mean)"
    ) +
    ggplot2::scale_color_continuous(type = "viridis", name = "Iteration") +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      legend.position = ifelse(plot_legend, "top", "none"),
      axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
      axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
      axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
      axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold")
    )
  
}

params <- c("penalty", "under_ratio", "over_ratio")
for(i in seq_along(params)) {
  print(plot_parameter_tuning(params[i], ifelse(i == 1, TRUE, FALSE)))
}

```
:::

### Evaluation on test set
**`r get_metrics_description(glmnet_low$evaluation$metrics_test)`**
```{r glmnet low cm}
plot_cm(glmnet_low$evaluation, "test")
```	
### Most important variables per metagenre
```{r glmnet low var importance}	
broom::tidy(glmnet_low$model) |> 
  dplyr::filter(term != "(Intercept)") |> 
  dplyr::group_by(class) |> 
  dplyr::slice_max(
    order_by = abs(estimate),
    n = 5
  ) |>
  dplyr::ungroup() |> 
  dplyr::arrange(class, dplyr::desc(abs(estimate))) |> 
  dplyr::mutate(estimate = round(estimate, 2)) |>
  dplyr::select(
    Metagenre = class,
    Feature = term,
    Coefficient = estimate
  ) |>
  knitr::kable(
    col.names = c("Metagenre", "Feature", "Coefficient"),
    caption = "Top 5 most important features per metagenre according to the GLMNET classifier for low-detailed metagenre labels."
  ) |>
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:3,
    border_right = FALSE,
    border_left = FALSE, 
  ) |>
  kableExtra::row_spec(
    0,
    bold = TRUE,
  ) |>  
  kableExtra::collapse_rows(columns = 1, latex_hline = "major", valign = "top")

```

## High-detailed metagenre labels
### Tuning
#### Model Selection
```{r glmnet high tuning top graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **GLMNET classifier for high-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."

glmnet_high <- load_classification_model(
  "../models/classifier/glmnet",
  "glmnet_highres"
)


chosen_parameters <- c(
  penalty = glmnet_high$settings$penalty,
  under_ratio = glmnet_high$settings$under_ratio, 
  over_ratio = glmnet_high$settings$over_ratio
)
# find matching model in tuning history
tuning_topmodels <- glmnet_high$tuning_history |> 
  dplyr::mutate(
    is_chosen = dplyr::if_else(
      penalty == chosen_parameters["penalty"] &
        under_ratio == chosen_parameters["under_ratio"] &
        over_ratio == chosen_parameters["over_ratio"],
      TRUE,
      FALSE
    )
  ) |> 
  dplyr::arrange(dplyr::desc(macro_f1_with_zeros_mean)) |> 
  head(top_n) |> 
  dplyr::mutate(rank = dplyr::row_number())
best_f1 <- tuning_topmodels$macro_f1_with_zeros_mean[1]
best_f1_se <- tuning_topmodels$macro_f1_with_zeros_std_err[1]
tuning_topmodels <-tuning_topmodels |> 
  dplyr::mutate(
    within_se_rule = macro_f1_with_zeros_mean >= (best_f1 - best_f1_se), 
    modelgroup = dplyr::case_when(
      is_chosen ~ 1,
      within_se_rule ~ 2,
      TRUE ~ 3
    ) |> as.factor()
  )


ggplot2::ggplot(tuning_topmodels, ggplot2::aes(x = rank, y = macro_f1_with_zeros_mean, color = modelgroup)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = macro_f1_with_zeros_mean - macro_f1_with_zeros_std_err,
      ymax = macro_f1_with_zeros_mean + macro_f1_with_zeros_std_err
    ),
    width = 0.2
  ) +
  ggplot2::labs(
    x = "Rank",
    y = "F1 Macro (mean ± se)"
  ) +
  ggplot2::scale_color_manual(
    values = c("#23cf00ff", "grey25", "grey75"),
  ) +
  ggplot2::scale_x_discrete(position = "top", limits = factor(1:top_n)) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.position = "none", 
    axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold"),
    panel.grid.major.x = ggplot2::element_blank()
  )



top_table <- tuning_topmodels |> 
  dplyr::mutate(
    `F1 Macro (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      macro_f1_with_zeros_mean,
      macro_f1_with_zeros_std_err
    ), 
    `Kappa (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      kap_mean,
      kap_std_err
    ), 
    `Accuracy (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      accuracy_mean,
      accuracy_std_err
    ), 
    `Undersampling Ratio` = round(under_ratio, 2),
    `Oversampling Ratio` = round(over_ratio, 2),
    Penalty = format(penalty, scientific = TRUE, digits = 3)
  ) |> 
    dplyr::select(
      Rank = rank,
      Iteration = .iter,
      Penalty,
      `Oversampling Ratio`,
      `Undersampling Ratio`,
      `F1 Macro (mean ± se)`,
      `Kappa (mean ± se)`,
      `Accuracy (mean ± se)`, 
      modelgroup
    )
```	
```{r glmnet high tuning top table}
knitr::kable(
  top_table,
  col.names = c(
    "Rank",
    "Iteration",
    "Penalty",
    "Oversampling Ratio",
    "Undersampling Ratio",
    "F1 Macro (mean ± se)",
    "Kappa (mean ± se)",
    "Accuracy (mean ± se)",
    ""
  )
) |> 
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:8,
    border_right = FALSE,
    border_left = FALSE, 
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 1),
    bold = TRUE,
    background = "#23cf00ff"
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 2),
    background = "grey90"
  ) |> 
  kableExtra::row_spec(
   which(top_table$modelgroup == 3),
    color = "grey35"
  ) |>
  kableExtra::remove_column(9)
 
```	
#### Parameter Exploration
:::{.column-screen-inset-right}
```{r glmnet high tuning param graphs}
#| layout-ncol: 2
make_axis_title <- function(x) {
  with_space <- gsub("[_\\.]", " ", x)
  split <- strsplit(with_space, " ")[[1]]
  capitalized <- paste(toupper(substring(split, 1, 1)), substring(split, 2),
        sep = "", collapse = " ")
  capitalized
}

plot_parameter_tuning <- function(parameter, plot_legend) {
  ggplot2::ggplot(glmnet_high$tuning_history, ggplot2::aes(x = .data[[parameter]], y = macro_f1_with_zeros_mean, color = .iter)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::scale_x_continuous(position = "top") +
    ggplot2::labs(
      x = make_axis_title(parameter),
      y = "F1 Macro (mean)"
    ) +
    ggplot2::scale_color_continuous(type = "viridis", name = "Iteration") +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      legend.position = ifelse(plot_legend, "top", "none"),
      axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
      axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
      axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
      axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold")
    )
  
}

params <- c("penalty", "under_ratio", "over_ratio")
for(i in seq_along(params)) {
  print(plot_parameter_tuning(params[i], ifelse(i == 1, TRUE, FALSE)))
}

```
:::

### Evaluation on test set
**`r get_metrics_description(glmnet_high$evaluation$metrics_test)`**
```{r glmnet high cm, fig.width=10, fig.height=8.3}
plot_cm(glmnet_high$evaluation, "test")
```	
### Most important variables per metagenre
```{r glmnet high var importance}	
broom::tidy(glmnet_high$model) |> 
  dplyr::filter(term != "(Intercept)") |> 
  dplyr::group_by(class) |> 
  dplyr::slice_max(
    order_by = abs(estimate),
    n = n_glmnet_varimp
  ) |>
  dplyr::ungroup() |> 
  dplyr::arrange(class, dplyr::desc(abs(estimate))) |> 
  dplyr::mutate(estimate = round(estimate, 2)) |>
  dplyr::select(
    Metagenre = class,
    Feature = term,
    Coefficient = estimate
  ) |>
  knitr::kable(
    col.names = c("Metagenre", "Feature", "Coefficient"),
    caption = "Most important features per metagenre according to the GLMNET classifier for high-detailed metagenre labels."
  ) |>
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:3,
    border_right = FALSE,
    border_left = FALSE, 
  ) |>
  kableExtra::row_spec(
    0,
    bold = TRUE,
  ) |>  
  kableExtra::collapse_rows(columns = 1, latex_hline = "major", valign = "top")

```

# Random Forest Classifier
## Low-detailed metagenre labels
### Tuning
#### Model Selection
```{r rf low tuning top graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **Random Forest classifier for low-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."

rf_low <- list()
rf_low$settings <- readRDS("../models/classifier/rf/rf_highres_settings.rds")
rf_low$evaluation <- readRDS("../models/classifier/rf/rf_highres_evaluation.rds")
rf_low$tuning_history <- readRDS("../models/classifier/rf/rf_highres_tuning_history.rds")


chosen_parameters <- c(
  max.depth = rf_low$settings$max.depth,
  min_n = rf_low$settings$min_n,
  mtry = rf_low$settings$mtry,
  under_ratio = rf_low$settings$under_ratio, 
  over_ratio = rf_low$settings$over_ratio
)
# find matching model in tuning history
tuning_topmodels <- rf_low$tuning_history |> 
  dplyr::mutate(
    is_chosen = dplyr::if_else(
      max.depth == chosen_parameters["max.depth"] &
        min_n == chosen_parameters["min_n"] &
        mtry == chosen_parameters["mtry"] &
        under_ratio == chosen_parameters["under_ratio"] &
        over_ratio == chosen_parameters["over_ratio"],
      TRUE,
      FALSE
    )
  ) |> 
  dplyr::arrange(dplyr::desc(macro_f1_with_zeros_mean)) |> 
  head(top_n) |> 
  dplyr::mutate(rank = dplyr::row_number())
best_f1 <- tuning_topmodels$macro_f1_with_zeros_mean[1]
best_f1_se <- tuning_topmodels$macro_f1_with_zeros_std_err[1]
tuning_topmodels <-tuning_topmodels |> 
  dplyr::mutate(
    within_se_rule = macro_f1_with_zeros_mean >= (best_f1 - best_f1_se), 
    modelgroup = dplyr::case_when(
      is_chosen ~ 1,
      within_se_rule ~ 2,
      TRUE ~ 3
    ) |> as.factor()
  )


ggplot2::ggplot(tuning_topmodels, ggplot2::aes(x = rank, y = macro_f1_with_zeros_mean, color = modelgroup)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = macro_f1_with_zeros_mean - macro_f1_with_zeros_std_err,
      ymax = macro_f1_with_zeros_mean + macro_f1_with_zeros_std_err
    ),
    width = 0.2
  ) +
  ggplot2::labs(
    x = "Rank",
    y = "F1 Macro (mean ± se)"
  ) +
  ggplot2::scale_color_manual(
    values = c("#23cf00ff", "grey25", "grey75"),
  ) +
  ggplot2::scale_x_discrete(position = "top", limits = factor(1:top_n)) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.position = "none", 
    axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold"),
    panel.grid.major.x = ggplot2::element_blank()
  )



top_table <- tuning_topmodels |> 
  dplyr::mutate(
    `F1 Macro (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      macro_f1_with_zeros_mean,
      macro_f1_with_zeros_std_err
    ), 
    `Kappa (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      kap_mean,
      kap_std_err
    ), 
    `Accuracy (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      accuracy_mean,
      accuracy_std_err
    ), 
    `Undersampling Ratio` = round(under_ratio, 2),
    `Oversampling Ratio` = round(over_ratio, 2)
  ) |> 
    dplyr::select(
      Rank = rank,
      Iteration = .iter,
      `Max Depth` = max.depth,
      `Min Node Size` = min_n,
      `MTRY` = mtry,
      `Oversampling Ratio`,
      `Undersampling Ratio`,
      `F1 Macro (mean ± se)`,
      `Kappa (mean ± se)`,
      `Accuracy (mean ± se)`, 
      modelgroup
    )
```	
```{r rf low tuning top table}
knitr::kable(
  top_table,
  col.names = c(
    "Rank",
    "Iteration",
    "Max Depth",
    "Min Node Size",
    "MTRY",
    "Oversampling Ratio",
    "Undersampling Ratio",
    "F1 Macro (mean ± se)",
    "Kappa (mean ± se)",
    "Accuracy (mean ± se)",
    ""
  )
) |> 
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:8,
    border_right = FALSE,
    border_left = FALSE, 
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 1),
    bold = TRUE,
    background = "#23cf00ff"
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 2),
    background = "grey90"
  ) |> 
  kableExtra::row_spec(
   which(top_table$modelgroup == 3),
    color = "grey35"
  ) |>
  kableExtra::remove_column(9)
 
```	
#### Parameter Exploration
:::{.column-screen-inset-right}
```{r rf low tuning param graphs}
#| layout-ncol: 2
make_axis_title <- function(x) {
  with_space <- gsub("[_\\.]", " ", x)
  split <- strsplit(with_space, " ")[[1]]
  capitalized <- paste(toupper(substring(split, 1, 1)), substring(split, 2),
        sep = "", collapse = " ")
  capitalized
}

plot_parameter_tuning <- function(parameter, plot_legend) {
  ggplot2::ggplot(rf_low$tuning_history, ggplot2::aes(x = .data[[parameter]], y = macro_f1_with_zeros_mean, color = .iter)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::scale_x_continuous(position = "top") +
    ggplot2::labs(
      x = make_axis_title(parameter),
      y = "F1 Macro (mean)"
    ) +
    ggplot2::scale_color_continuous(type = "viridis", name = "Iteration") +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      legend.position = ifelse(plot_legend, "top", "none"),
      axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
      axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
      axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
      axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold")
    )
  
}

params <- c("max.depth", "min_n", "mtry", "under_ratio", "over_ratio")
for(i in seq_along(params)) {
  print(plot_parameter_tuning(params[i], ifelse(i == 1, TRUE, FALSE)))
}

```
:::

### Evaluation on test set
**`r get_metrics_description(rf_low$evaluation$metrics_test)`**
```{r rf low cm}
plot_cm(rf_low$evaluation, "test")
```	
### Most important variables per metagenre
```{r rf low var importance}	
rf_low$evaluation$varimp |> 
  head(n_rf_varimp) |> 
  dplyr::arrange(dplyr::desc(Importance)) |>
  dplyr::mutate(
    Variable = forcats::fct_inorder(Variable) |> forcats::fct_rev()
  ) |>
  ggplot2::ggplot(ggplot2::aes(y = Variable, x = Importance)) +	
  ggplot2::geom_bar(stat = "identity", fill = "grey65") + 
  ggplot2::scale_x_continuous(position = "top", expand = ggplot2::expansion(mult = c(0, 0.05))) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.title.y = ggplot2::element_blank(),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 16, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12), 
    panel.grid.major.y = ggplot2::element_blank()
  )
```

## High-detailed metagenre labels
### Tuning
#### Model Selection
```{r rf high tuning top graph}
#| fig-cap: "**Best hyperparameter configurations** from tuning of the **Random Forest classifier for high-detailed** metagenre labels. The chosen hyperparameter configuration is highlighted in green, other **models within one standard error of the numerically best model** are shown in dark grey."

rf_high <- list()
rf_high$settings <- readRDS("../models/classifier/rf/rf_highres_settings.rds")
rf_high$evaluation <- readRDS("../models/classifier/rf/rf_highres_evaluation.rds")
rf_high$tuning_history <- readRDS("../models/classifier/rf/rf_highres_tuning_history.rds")


chosen_parameters <- c(
  max.depth = rf_high$settings$max.depth,
  min_n = rf_high$settings$min_n,
  mtry = rf_high$settings$mtry,
  under_ratio = rf_high$settings$under_ratio, 
  over_ratio = rf_high$settings$over_ratio
)
# find matching model in tuning history
tuning_topmodels <- rf_high$tuning_history |> 
  dplyr::mutate(
    is_chosen = dplyr::if_else(
      max.depth == chosen_parameters["max.depth"] &
        min_n == chosen_parameters["min_n"] &
        mtry == chosen_parameters["mtry"] &
        under_ratio == chosen_parameters["under_ratio"] &
        over_ratio == chosen_parameters["over_ratio"],
      TRUE,
      FALSE
    )
  ) |> 
  dplyr::arrange(dplyr::desc(macro_f1_with_zeros_mean)) |> 
  head(top_n) |> 
  dplyr::mutate(rank = dplyr::row_number())
best_f1 <- tuning_topmodels$macro_f1_with_zeros_mean[1]
best_f1_se <- tuning_topmodels$macro_f1_with_zeros_std_err[1]
tuning_topmodels <-tuning_topmodels |> 
  dplyr::mutate(
    within_se_rule = macro_f1_with_zeros_mean >= (best_f1 - best_f1_se), 
    modelgroup = dplyr::case_when(
      is_chosen ~ 1,
      within_se_rule ~ 2,
      TRUE ~ 3
    ) |> as.factor()
  )


ggplot2::ggplot(tuning_topmodels, ggplot2::aes(x = rank, y = macro_f1_with_zeros_mean, color = modelgroup)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = macro_f1_with_zeros_mean - macro_f1_with_zeros_std_err,
      ymax = macro_f1_with_zeros_mean + macro_f1_with_zeros_std_err
    ),
    width = 0.2
  ) +
  ggplot2::labs(
    x = "Rank",
    y = "F1 Macro (mean ± se)"
  ) +
  ggplot2::scale_color_manual(
    values = c("#23cf00ff", "grey25", "grey75"),
  ) +
  ggplot2::scale_x_discrete(position = "top", limits = factor(1:top_n)) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.position = "none", 
    axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold"),
    panel.grid.major.x = ggplot2::element_blank()
  )



top_table <- tuning_topmodels |> 
  dplyr::mutate(
    `F1 Macro (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      macro_f1_with_zeros_mean,
      macro_f1_with_zeros_std_err
    ), 
    `Kappa (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      kap_mean,
      kap_std_err
    ), 
    `Accuracy (mean ± se)` = sprintf(
      "%.3f ± %.3f",
      accuracy_mean,
      accuracy_std_err
    ), 
    `Undersampling Ratio` = round(under_ratio, 2),
    `Oversampling Ratio` = round(over_ratio, 2)
  ) |> 
    dplyr::select(
      Rank = rank,
      Iteration = .iter,
      `Max Depth` = max.depth,
      `Min Node Size` = min_n,
      `MTRY` = mtry,
      `Oversampling Ratio`,
      `Undersampling Ratio`,
      `F1 Macro (mean ± se)`,
      `Kappa (mean ± se)`,
      `Accuracy (mean ± se)`, 
      modelgroup
    )
```	
```{r rf high tuning top table}
knitr::kable(
  top_table,
  col.names = c(
    "Rank",
    "Iteration",
    "Max Depth",
    "Min Node Size",
    "MTRY",
    "Oversampling Ratio",
    "Undersampling Ratio",
    "F1 Macro (mean ± se)",
    "Kappa (mean ± se)",
    "Accuracy (mean ± se)",
    ""
  )
) |> 
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
  ) |>
  kableExtra::column_spec(
    1:8,
    border_right = FALSE,
    border_left = FALSE, 
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 1),
    bold = TRUE,
    background = "#23cf00ff"
  ) |> 
  kableExtra::row_spec(
    which(top_table$modelgroup == 2),
    background = "grey90"
  ) |> 
  kableExtra::row_spec(
   which(top_table$modelgroup == 3),
    color = "grey35"
  ) |>
  kableExtra::remove_column(9)
 
```	
#### Parameter Exploration
:::{.column-screen-inset-right}
```{r rf high tuning param graphs}
#| layout-ncol: 2
make_axis_title <- function(x) {
  with_space <- gsub("[_\\.]", " ", x)
  split <- strsplit(with_space, " ")[[1]]
  capitalized <- paste(toupper(substring(split, 1, 1)), substring(split, 2),
        sep = "", collapse = " ")
  capitalized
}

plot_parameter_tuning <- function(parameter, plot_legend) {
  ggplot2::ggplot(rf_high$tuning_history, ggplot2::aes(x = .data[[parameter]], y = macro_f1_with_zeros_mean, color = .iter)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::scale_x_continuous(position = "top") +
    ggplot2::labs(
      x = make_axis_title(parameter),
      y = "F1 Macro (mean)"
    ) +
    ggplot2::scale_color_continuous(type = "viridis", name = "Iteration") +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      legend.position = ifelse(plot_legend, "top", "none"),
      axis.title.y = ggplot2::element_text(color = "grey45", size = 14, hjust = 1),
      axis.title.x = ggplot2::element_text(color = "grey45", size = 14, hjust = 0),
      axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
      axis.text.y = ggplot2::element_text(color = "black", size = 12, face = "bold")
    )
  
}

params <- c("max.depth", "min_n", "mtry", "under_ratio", "over_ratio")
for(i in seq_along(params)) {
  print(plot_parameter_tuning(params[i], ifelse(i == 1, TRUE, FALSE)))
}

```
:::

### Evaluation on test set
**`r get_metrics_description(rf_high$evaluation$metrics_test)`**
```{r rf high cm, fig.width=10, fig.height=8.3}
plot_cm(rf_high$evaluation, "test")
```	
### Most important variables per metagenre
```{r rf high var importance}	
rf_high$evaluation$varimp |> 
  head(n_rf_varimp) |> 
  dplyr::arrange(dplyr::desc(Importance)) |>
  dplyr::mutate(
    Variable = forcats::fct_inorder(Variable) |> forcats::fct_rev()
  ) |>
  ggplot2::ggplot(ggplot2::aes(y = Variable, x = Importance)) +	
  ggplot2::geom_bar(stat = "identity", fill = "grey65") + 
  ggplot2::scale_x_continuous(position = "top", expand = ggplot2::expansion(mult = c(0, 0.05))) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.title.y = ggplot2::element_blank(),
    axis.title.x = ggplot2::element_text(color = "grey45", size = 16, hjust = 0),
    axis.text.x = ggplot2::element_text(color = "grey45", size = 12),
    axis.text.y = ggplot2::element_text(color = "black", size = 12), 
    panel.grid.major.y = ggplot2::element_blank()
  )
```