---
title: "Metagenre Classification - Genre Discourse Analysis"
author: "Markus Radke"
date: "`r Sys.Date()`"
format: 
    html: 
        mermaid: 
          theme: neutral
        code-fold: true
        code-summary: "Show/Hide Code"
        output-dir: reports
        toc: true
        toc-location: left
        toc-depth: 5
        toc-expand: 5
        number-sections: true
        theme: cosmo
        code-links:
        - text: Github Repository
          href: "https://github.com/markusradke/GenreDiscourseAnalysis"
        css: 
          styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()

data_settings <- readRDS("../models/classifier/data_settings.rds")

train_low <- readRDS("../models/classifier/train_low.rds")
test_low <- readRDS("../models/classifier/test_low.rds")
cv_splits_low <- readRDS("../models/classifier/cv_splits_low.rds")

train_medium <- readRDS("../models/classifier/train_medium.rds")
test_medium <- readRDS("../models/classifier/test_medium.rds")
cv_splits_medium <- readRDS("../models/classifier/cv_splits_medium.rds")

train_high <- readRDS("../models/classifier/train_high.rds")
test_high <- readRDS("../models/classifier/test_high.rds")
cv_splits_high <- readRDS("../models/classifier/cv_splits_high.rds")

train_very_high <- readRDS("../models/classifier/train_very_high.rds")
test_very_high <- readRDS("../models/classifier/test_very_high.rds")
cv_splits_very_high <- readRDS("../models/classifier/cv_splits_very_high.rds")

top_n <- 40 # top n models to show in report
n_top_varimp <- 20

lowres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_low.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max)) 

mediumres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_medium.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max))

highres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_high.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max))

veryhighres_certainty <- read_feather_with_lists("../models/metagenres/mb_metagenres_very_high.feather") |> 
  dplyr::group_by(metagenre) |> 
  dplyr::summarize(certainty = median(p_max))
```

# Complete Workflow
:::{.column-screen-inset-right}
```{mermaid}
%%{init: {
  'flowchart': {
    'useMaxWidth': true,
    'htmlLabels': false
  }
}}%%

flowchart TB
 subgraph Data["Data prep"]
    direction LR
        import["import POPTRAG"]
        select["select potential features"]
        transform["transform chars to factors </br> apply lyrics heuristics"]
        casewise["drop observations with more than 40% missing"]
        target["join metagenre </br>target labels"]
        minn["drop levels with less than 100 observations in one of the splits / folds"]
        dummy["dummy code categorical predictors</br>reference level: most common"]
  end
 subgraph Split["Spliting Strategy"]
    direction LR
        testtrain["split train/test <br> (0.8/0.2, artist-level)"]
        cvfold["create 10-fold CV splits </br>(artist-level)"]
  end
 subgraph Pre["Preprocessing"]
    direction LR
        impute["impute median"]
        down["nearmiss downsample majority classes </br>(factor tuned)"]
        up["upsample minority </br>(broderline SMOTE, factor tuned)"]
        rmzv["remove zero-variance predictors"]
        normalize["normalize numeric </br>(z-score)"]
  end
 subgraph Model["Modeling and eval"]
    direction LR
        tune["tune hyperparams on F1 macro </br>(Bayesian optimization, </br>20-set latin hypercube inital grid,</br> max 50 iters, 10-iters stop)"]
        best["select best params </br>(based on F1 macro, one standard error rule)"]
        finalfit["fit final model </br>(on full training set)"]
        evaluate["evaluate on test set"]
  end
    import --> select
    select --> transform
    transform --> casewise
    casewise --> target
    target --> minn
    minn --> dummy
    minn <-.-> Split

    testtrain --> cvfold

    impute -. only during training .-> down
    impute --> rmzv
    down -.-> up
    up -.-> rmzv
    rmzv --> normalize
    tune --> best
    tune <--> Pre
    best --> finalfit
    finalfit --> evaluate
    evaluate <--> Pre
    dummy --> Split
    Split --> Model

    style Pre fill:#f8f0e8,stroke:#cc6600,stroke-width:3px,color:#000
    style Data fill:#e8f4f8,stroke:#0066cc,stroke-width:3px,color:#000
    style Split fill:#f0e8f8,stroke:#6600cc,stroke-width:3px,color:#000
    style Model fill:#e8f8e8,stroke:#00aa00,stroke-width:3px,color:#000
  
```
:::

:::{.callout-note}
For up- and down-sampling, we employed adaptive balancing: We employed a target-ratio parameter (tunable range: 1 to observed max:min ratio) that governed the degree of class rebalancing. The algorithm computed a reference count from class frequencies (median for <= 10 classes; 60th percentile for 11-20 classes; 65th percentile for > 20 classes), then downsampled classes exceeding reference$\times$target ratio using NearMiss and upsampled classes below reference < target ratio using Borderline SMOTE, and regular SMOTE in case of non sufficient danger observations.
:::


:::{.callout-note}
Random forests were tuned with 500 trees per ensemble during tuning. The final model on the whole tranining set was trained with 1000 trees.
:::

# Data Summary

```{r dataset_summary}
summarize_ds <- function(resolution, split, df) {
  data.frame(
    Metagenre_Resolution = resolution,
    Split = split,
    N_Tracks = format(nrow(df), big.mark = ","),
    N_Artists = format(nrow(dplyr::distinct(df, .data$artist.s.id)), big.mark = ",")
  )
}

datasets <- dplyr::bind_rows(
  summarize_ds("low (10-15)",  "Train", train_low),
  summarize_ds("low (10-15)",  "Test",  test_low),
  summarize_ds("high (25-30)", "Train", train_high),
  summarize_ds("high (25-30)", "Test",  test_high)
)

knitr::kable(
  datasets,
  col.names = c("Metagenre Resolution", "Split", "N~Tracks~", "N~Artists~"),
  caption = "Dataset splits used in the analyses"
)
```

# Split outcome distributions

```{r outcome_distributions_low}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **low-detailed** meta-genre taxonomy."
distribution_low <- compare_metagenre_distributions_splits(train_low, test_low, cv_splits_low)
distribution_low$plot
```

```{r outcome_distributions_medium}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **high-detailed** detailed meta-genre taxonomy."
distribution_medium <- compare_metagenre_distributions_splits(train_medium, test_medium, cv_splits_medium)
distribution_medium$plot
```

```{r outcome_distributions_high}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **high-detailed** detailed meta-genre taxonomy."
distribution_high <- compare_metagenre_distributions_splits(train_high, test_high, cv_splits_high)
distribution_high$plot
```

```{r outcome_distributions_very_high}
#| fig-cap: "Comparison of metagenre distributions in train and test splits for **high-detailed** detailed meta-genre taxonomy."
distribution_very_high <- compare_metagenre_distributions_splits(
  train_very_high,
  test_very_high, 
  cv_splits_very_high
)
distribution_very_high$plot
```

# Model Features

:::{.column-screen-inset-right}
```{r model_features}
model_features <- colnames(train_low)[
  !colnames(train_low) %in% c("track.s.id", "artist.s.id", "metagenre", "case_wts")
] |> sort()
```


A total of `r length(model_features)` features was used for training the classifiers on the full data set.

:::{.callout-note}
For training on a subsample of the dataset the number of model features was reduced to 171 due to the requirment of 100 observations per class and fold in every factor level. 
:::

```{r model_features_table}
model_features_padded <- c(
  model_features,
  rep("", times = 5 - (length(model_features) %% 5))
)

knitr::kable(
  matrix(model_features_padded, ncol = 5, byrow = TRUE),
  col.names = c("Model features", rep("", 4))
) |>
  kableExtra::kable_styling(full_width = FALSE, position = "left") |>
  kableExtra::column_spec(1:5, border_right = FALSE, border_left = FALSE) |>
  kableExtra::row_spec(2:ceiling(length(model_features_padded) / 5), hline_after = FALSE)
```
:::

:::{.callout-note}
Variables starting with "dtb." refer to artist genres from Spotify. We assume that artist genres are assigned with a strategic aim for distribution on the platform (Valverde, 2025). In that regard, they are different from the musical genres assigned by the listeners. The variables are flags indicating whether an artist genre is assigned to a track's artist or not. To reduce the dimensionality, we used the same hiearchical-tree-based approach as for the MusicBrainz meta-genres.
:::

# Choosing the best learner on 20% subsample
```{r best_learner_comparison}
load_learner_f1_cv <- function(learner, resolution, hash, model_dir, param_names) {
  tuning <- readRDS(
    sprintf(
      "../models/classifier/%s/%s_%s_%s_tuning_history.rds",
      model_dir, learner, resolution, hash
    )
  )
  settings <- readRDS(
    sprintf(
      "../models/classifier/%s/%s_%s_%s_settings.rds",
      model_dir, learner, resolution, hash
    )
  )
  params <- unlist(param_names)
  chosen_params <- setNames(unlist(settings[params]), params)
  # exchange "alpha" with "mixture" if in names for GLMNET
  if (learner == "GLMNET"){
    names(chosen_params) <- gsub("alpha", "mixture", names(chosen_params))
  }
   
  prepare_tuning_topmodels(tuning, chosen_params, top_n = 100) |>
    dplyr::filter(is_chosen) |>
    dplyr::select(
      macro_f1_with_zeros_mean,
      macro_f1_with_zeros_std_err
    ) |>
    dplyr::mutate(learner = learner, detail = resolution)
}

model_configs <- tibble::tribble(
  ~learner, ~resolution, ~hash, ~model_dir, ~param_names,
  "GLMNET", "lowres", "436e205a", "glmnet", 
    list(c("penalty", "alpha", "target_ratio")),
  "GLMNET", "mediumres", "01f3f5dd", "glmnet", 
    list(c("penalty", "alpha", "target_ratio")),
  "RDA", "lowres", "d07f8bb4", "rda", 
    list(c("gamma", "lambda", "target_ratio")),
  "RDA", "mediumres", "4d480cd8", "rda", 
    list(c("gamma", "lambda", "target_ratio")),
  "RF", "lowres", "759e2621", "rf", 
    list(c("mtry", "min_n", "max.depth", "target_ratio")),
  "RF", "mediumres", "2cc36c89", "rf", 
    list(c("mtry", "min_n", "max.depth", "target_ratio")),
  "LightGBM", "lowres", "1776db28", "lightgbm", 
    list(c("tree_depth", "trees", "learn_rate", "mtry", 
           "min_n", "target_ratio")),
  "LightGBM", "mediumres", "b8c05240", "lightgbm", 
    list(c("tree_depth", "trees", "learn_rate", "mtry", 
           "min_n", "target_ratio"))
)

chosen_models <- purrr::pmap_df(model_configs, load_learner_f1_cv)
learner_comp_plot <- plot_learner_comparison(chosen_models)

ggplot2::ggsave(
  "../reports/paper_v1/figures/learner_comparison.png",
  plot = learner_comp_plot,
  width = 180,
  height = 115,
  units = "mm",
  dpi = 300
)
learner_comp_plot
```


# Final GBM Classifier Results
```{r setup_resolutions}
resolutions <- list(
  low = list(
    hash = "dc8426d1",
    certainty = lowres_certainty,
    title = "Low-detailed"
  ),
  medium = list(
    hash = "811e1b4e",
    certainty = mediumres_certainty,
    title = "Medium-detailed"
  ),
  high = list(
    hash = "de549d7b",
    certainty = highres_certainty,
    title = "High-detailed"
  ),
  veryhigh = list(
    hash = "8b88c70b",
    certainty = veryhighres_certainty,
    title = "Very high-detailed"
  )
)

params <- c("tree_depth", "trees", "learn_rate", "mtry", "min_n", "target_ratio")

load_resolution_data <- function(resolution_hash) {
  base_path <- sprintf(
    "../models/classifier/lightgbm/final_models/lightgbm_%sres_%s",
    names(resolutions)[which(sapply(resolutions, `[[`, "hash") == resolution_hash)],
    resolution_hash
  )
  
  list(
    eval = readRDS(sprintf("%s_evaluation.rds", base_path)),
    settings = readRDS(sprintf("%s_settings.rds", base_path)),
    tuning = readRDS(sprintf("%s_tuning_history.rds", base_path))
  )
}

extract_chosen_params <- function(settings) {
  c(
    tree_depth = settings$tree_depth,
    trees = settings$trees,
    learn_rate = settings$learn_rate,
    mtry = settings$mtry,
    min_n = settings$min_n,
    target_ratio = settings$target_ratio
  )
}
```

# Final GBM Classifier Results

:::{.column-screen-inset-right}
```{r results, results='asis'}
for (res_name in names(resolutions)) {
  res_config <- resolutions[[res_name]]
  
  cat(sprintf("\n## %s supergenre labels\n\n", res_config$title))
  
  data <- load_resolution_data(res_config$hash)
  chosen <- extract_chosen_params(data$settings)
  topmodels <- prepare_tuning_topmodels(data$tuning, chosen, top_n)
  
  cat("### Tuning\n\n")
  cat("#### Model Selection\n\n")
  
  print(plot_tuning_model_selection(topmodels, top_n))
  
  table <- create_lgbm_top_table(topmodels, param_names = params)
  print(render_tuning_top_table(table))
  
  cat("\n#### Parameter Exploration\n\n")
  for (i in seq_along(params)) {
    print(plot_parameter_tuning(data$tuning, params[i], i == 1))
  }
  
  cat("### Evaluation on test set\n\n")
  
  cat(sprintf("**%s**\n\n", get_metrics_description(data$eval$metrics_test)))
  
  f1_scores <- calculate_class_f1_scores_from_confusion_df(data$eval$confusion_test)
  correlation <- round(cor(res_config$certainty$certainty, f1_scores), 3)
  cat(sprintf(
    "**Correlation between metagenre uncertainty and per-class F1 scores on test set: %s**\n\n",
    correlation
  ))
  
  cm_plot <- plot_cm(data$eval, res_config$certainty, "test")
  
  width_cm_paper <- ifelse(res_name == "low", 180, 360)
  width_cm_paper <- ifelse(res_name == "medium", 240, width_cm_paper)
  width_cm_paper <- ifelse(res_name == "high", 320, width_cm_paper)
  heigth_cm_paper <- ifelse(res_name == "low", 100, 225)
  heigth_cm_paper <- ifelse(res_name == "medium", 150, heigth_cm_paper)
  heigth_cm_paper <- ifelse(res_name == "high", 200, heigth_cm_paper)

  ggplot2::ggsave(
    cm_plot,
    filename = sprintf("../reports/paper_v1/figures/lightgbm_%sres_cm.png", res_name),
    width = width_cm_paper,
    height = heigth_cm_paper,
    units = "mm",
    dpi = 300
  )
  
  print(cm_plot)


  cat("\n#### Feature Importance\n\n")
  varimp_plot <- plot_gbm_varimp(
    data$eval$varimp,
    n_top = n_top_varimp
  )
  ggplot2::ggsave(
    varimp_plot,
    filename = sprintf("../reports/paper_v1/figures/lightgbm_%sres_varimp.png", res_name),
    width = 180,
    height = 150,
    units = "mm",
    dpi = 300
  )
  print(varimp_plot)
}
```
:::